Portable Multi- and Many-Core Performance for Finite Difference Codes; Application to the Free-Surface Component of NEMO.

### 2017-09-13 docs/papers/nemolite2d_performance/nemolite2d_perf.tex
nemolite2d-paper: WIP on new, Related Approaches section

The {PS}y{KA}l approach attempts to address the problems described in
the previous section. It separates code into three layers; the
Algorithm layer, the PSy layer and the Kernel layer. The approach has
been developed in the GungHo project~\citep{GungHo}, which is creating
a new Dynamical core for the UK Met Office, and its design has been
influenced by earlier work on OP2~\citep{OP2, PYOP2}.

In common with OP2, the {PS}y{KA}l approach separates out the science
code and the performance-related code into distinct layers. The calls
that specify parallelism in both approaches are similar in terms of
where they are placed in the code and in their semantics. However, the
{PS}y{KA}l approach supports the specification of more than one kernel
in a parallel region of code, compared with one for OP2, giving more
scope for optimisation. In addition, the metadata describing a kernel
is included with the kernel code in the {PS}y{KA}l approach whereas it
is provided as a part of the kernel call in OP2.

@@

The bottom layer, in terms of calling hierarchy, is the Kernel
layer. The Kernel layer implements the science that the Algorithm
layer calls, as a set of subroutines. These kernels operate on fields
that are local to the process doing the computation. (Depending on the
type of kernel, these may be a set of elements, a single column of
elements, or a set of columns.) Again the scientist is responsible for
writing this layer and there is no parallelism specified here, but,
depending on the complexity of the Kernels, there may be input from an
HPC expert and/or some coding rules to help ensure that the kernels
compile into efficient code.  In an alternative
approach~\citep{firedrake,fenics}, kernels are generated from a
high-level specification, potentially allowing them to be optimised
automatically~\citep{coffee}.

@@

>>>>

The {PS}y{KA}l approach attempts to address the problems described in
the previous section. It separates code into three layers; the
Algorithm layer, the PSy layer and the Kernel layer. The approach has
been developed in the GungHo project~\citep{GungHo}, which is creating
a new Dynamical core for the UK Met Office, and its design has been
influenced by earlier work on OP2~\citep{OP2, PYOP2} - see Section
\ref{sec_other_approaches} for more details.

@@

The bottom layer, in terms of calling hierarchy, is the Kernel
layer. The Kernel layer implements the science that the Algorithm
layer calls, as a set of subroutines. These kernels operate on fields
that are local to the process doing the computation. (Depending on the
type of kernel, these may be a set of elements, a single column of
elements, or a set of columns.) Again the scientist is responsible for
writing this layer and there is no parallelism specified here, but,
depending on the complexity of the Kernels, there may be input from an
HPC expert and/or some coding rules to help ensure that the kernels
compile into efficient code.

@@

\subsection{Related Approaches}
\label{sec_other_approaches}

As already mentioned, our approach is heavily influenced by the OP2
system~\citep{OP2, PYOP2}.
In common with OP2, the {PS}y{KA}l approach separates out the science
code and the performance-related code into distinct layers. The calls
that specify parallelism in both approaches are similar in terms of
where they are placed in the code and in their semantics. However, the
{PS}y{KA}l approach supports the specification of more than one kernel
in a parallel region of code, compared with one for OP2, giving more
scope for optimisation. In addition, the metadata describing a kernel
is included with the kernel code in the {PS}y{KA}l approach whereas it
is provided as a part of the kernel call in OP2.

Although {PS}y{KA}l is a design for architecting code, rather than an
implementation of a tool, there is an implicit assumption that the
majority of the kernels in an application will be provided by the
application developer. In the GridTools~\citep{grid_tools} and
Firedrake~\citep{firedrake,fenics} approaches, the mathematical
operations (finite-difference stencils and finite element operations,
respectively) are specified in a high-level language by the
user. Kernel code is then generated automatically. It is possible to
optimise this generated code~\citep{coffee} and, in the case of
GridTools, support both CPU and GPU architectures.

% Put this later after we've discussed porting nemolite2d to GPU?
%It is likely that the {PS}y{KA}l approach of constraining all
%performance-related code to be within the middle, {PS}y, layer

Lower-level abstractions, e.g. OCCA.

H. C. Edwards, C. R. Trott, D. Sunderland, "Kokkos: Enabling manycore performance portability through polymorphic memory access patterns", J. Parallel Distrib. Comput., vol. 74, no. 12, pp. 3202-3216, 2014.
